Appendix to ``Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis''

### 2016-12-03 MHCpaper/MHCsupp.tex
sentence about long term
main figure references
surya's edits

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[10pt]{article}
%\usepackage[printfigures]{figcaps}
%\usepackage[nolists,nomarkers]{endfloat}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
%\newcounter{tablerow}[table]
%\newcommand{\tablerowautorefname}{row}
%\newenvironment{tabularn}[1]{\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
\renewcommand{\figurename}{Appendix-figure}
\renewcommand{\figureautorefname}{Appendix-figure}
\DeclareCaptionLabelFormat{supp}{\textbf{#1 #2}}
\captionsetup{labelformat=supp}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{K$^\mathrm{b}$D$^\mathrm{b}$-/-}}}
\newcommand{\KO}{K$^\mathrm{b}$D$^{\mathrm{b}-/-}$}
%\newcommand{\ko}{_{\text{MHC}}}
%\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
%\newcommand{\lmax}{_{\text{max}}}
%\newcommand{\lmin}{_{\text{min}}}
\newcommand{\modelfig}[1][A]{Figure 3#1 of the main text}
\newcommand{\datafig}[1][B]{Figure 2#1 of the main text}
\newcommand{\chrfig}[1][B]{Figure 4#1 of the main text}
\newcommand{\compfig}[1][A]{Figure 4 -- figure supplement 2#1}
\newcommand{\basefig}{Figure 1 -- figure supplement 2}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Appendix to ``Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis''}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}
\author{}
\date{}

@@

We make the following assumptions:
\begin{itemize}
  \item There synapses are identical and have $M$ internal functional states, where the states of different synapses are independent of each other.
%  \item The synapses that are eligible for plasticity are chosen randomly.
  \item Candidate potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $r\potdep$.
  \item Candidate potentiation and depression events are described by Markov processes with transition probabilities $\M\potdep$.
  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take values in a finite range that we can shift to $\pm1$. Most of the models will only use the two extreme values.
\end{itemize}

@@

For the serial, multistate and two state models, we will use the same value for the transition probabilities, $q$, for potentiation in the wild-type and \KO\ models.
We will use a larger value for $q$ for depression in the \KO\ models than in the wild-type.

For the cascade model, we will use the same value for the parameter $x$ (which controls the decay of transition rates, see \cite{Fusi2005cascade}) for potentiation in the wild-type and \KO\ models.
We will use a larger value for $x$ for depression in the \KO\ models than in the wild-type.

@@

Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed, as shown in \autoref{fig:models}\ref{fig:pooled_model}.
We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single compound synapse.

@@

This compound synapse would seem to have $2^P$ internal states.
However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
The transition network will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
%
\begin{equation}\label{eq:pooledweight}
  \w_i = \frac{2i}{P}-1.
\end{equation}
%

The Markov process is lumpable \wrt this partition of states (see \cite{kemeny1960finite,burke1958markovian,Ball1993Lumpability}).
The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$,
which must be the same for all states in lump $i$.

@@

The features of the experiments shown in \modelfig\ that we'd like to capture are:
%
\begin{enumerate}
  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the \KO\ mice.
  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
  \item For the \KO\ mice, gain-increase learning is significantly faster with pre-training than without it.
  \item After pre-training, gain-increase learning is significantly faster in the \KO\ mice than in the wild-type.
%  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
\end{enumerate}
%
These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
\emph{We will not study gain-decrease learning}, as its mechanisms are not fully understood, and known to be different to to gain-increase.
We are only modelling the effect of gain-decrease training on \emph{these} synapses.
We will also not discuss the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
%However, both the wild-type and \KO\ mice should be affected in the same way, so the difference in the curvature of the \KO\ mice and wild-type curves after pre-training would be nice to capture.

@@

The fact that after pre-training, gain-increase learning is significantly faster in the \KO\ models than in the wild-type, as discussed in the \hyperref[sec:results]{previous section}, is true of the initial learning rate.
With a long enough duration of training, memory of the initial conditions should be erased and the final learning outcome should be the same with or without pre-training.
In \autoref{fig:longterm} we show the result of longer training durations on the three models that were not ruled out on the basis of the initial learning rate, the serial, cascade and nonuniform multistate models.
For all of these models, we see that the wild-type models eventually catch up and overtake the \KO\.
The initial learning enhancement is a transient effect, and at later times it turns into a learning impairment.

@@

We see that, for all parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
The normal baseline oculomotor performance must have some other explanation, such as those outlined in \compfig.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[10pt]{article}
%\usepackage[printfigures]{figcaps}
%\usepackage[nolists,nomarkers]{endfloat}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
%\newcounter{tablerow}[table]
%\newcommand{\tablerowautorefname}{row}
%\newenvironment{tabularn}[1]{\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
\renewcommand{\figurename}{Appendix-figure}
\renewcommand{\figureautorefname}{Appendix-figure}
\DeclareCaptionLabelFormat{supp}{\textbf{#1 #2}}
\captionsetup{labelformat=supp}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{K$^\mathrm{b}$D$^\mathrm{b}$-/-}}}
\newcommand{\KO}{K$^\mathrm{b}$D$^{\mathrm{b}-/-}$}
%\newcommand{\ko}{_{\text{MHC}}}
%\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
%\newcommand{\lmax}{_{\text{max}}}
%\newcommand{\lmin}{_{\text{min}}}
\newcommand{\modelfig}[1][A]{Figure 4#1 of the main text}
\newcommand{\datafig}[1][B]{Figure 3#1 of the main text}
\newcommand{\chrfig}[1][B]{Figure 2#1 of the main text}
\newcommand{\compfig}[1][A]{Figure 2 -- figure supplement 2#1}
\newcommand{\basefig}{Figure 1 -- figure supplement 2}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Appendix to ``Understanding both enhanced and impaired learning with enhanced plasticity: a saturation hypothesis''}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}
\author{}
\date{}

@@

We make the following assumptions:
\begin{itemize}
  \item The synapses are identical and have $M$ internal functional states, where the states of different synapses are independent of each other.
%  \item The synapses that are eligible for plasticity are chosen randomly.
  \item Candidate potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $r\potdep$.
  \item Candidate potentiation and depression events are described by Markov processes with transition probabilities $\M\potdep$.
  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take values in a finite range that we can shift to $\pm1$. Most of the models will only use the two extreme values.
\end{itemize}

@@

For the serial, multistate and two state models, we will use the same value for $q$, the transition rate between adjacent states, for potentiation in the wild-type and \KO\ models.
We will use a larger value for $q$ for depression in the \KO\ models than in the wild-type.

For the cascade model, we will use the same value for the parameter $x$ (which controls the ratio of adjacent transition rates, see \cite{Fusi2005cascade}) for potentiation in the wild-type and \KO\ models.
We will use a larger value for $x$ for depression in the \KO\ models than in the wild-type.

@@

Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed and replenished when this is reversed, as shown in \autoref{fig:models}\ref{fig:pooled_model}.
We can avoid going beyond the independent synapse model by modelling the pool of synapses that share the resource as a single compound synapse.

@@

This compound synapse would seem to have $2^P$ internal states.
However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
We can group those states of the pool of synapses that have the same total number of potentiated synapses into a single state of the equivalent compound synapse.
The transition network will then have a multistate topology (see \autoref{fig:models}\ref{fig:multistate_model}) but the transition probabilities will no longer be uniform and the weight of the compound synapse is the mean of its constituent synapses:
%
\begin{equation}\label{eq:pooledweight}
  \w_i = \frac{2i}{P}-1.
\end{equation}
%

The Markov process is lumpable when partitioning states by total number of potentiated synapses (see \cite{kemeny1960finite,burke1958markovian,Ball1993Lumpability}).
The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$,
which must be the same for all states in lump $i$.

@@

The features of the experiments shown in \modelfig\ that we'd like to capture are:
%
\begin{enumerate}
  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the \KO\ mice.
  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
  \item For the \KO\ mice, gain-increase learning is significantly faster with pre-training than without it.
  \item After pre-training, gain-increase learning is significantly faster in the \KO\ mice than in the wild-type.
%  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
\end{enumerate}
%
These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
\emph{We will not study gain-decrease learning}, as its mechanisms are not fully understood, and known to be different to gain-increase.
We are only modelling the effect of gain-decrease training on \emph{these} synapses.
We will also not discuss the curvature of the learning curves, as this can be changed by the nonlinear relation between synaptic weight and VOR gain.
%However, both the wild-type and \KO\ mice should be affected in the same way, so the difference in the curvature of the \KO\ mice and wild-type curves after pre-training would be nice to capture.

@@

The fact that after pre-training, gain-increase learning is significantly faster in the \KO\ models than in the wild-type, as discussed in the \hyperref[sec:results]{previous section}, is true of the initial learning rate.
With a long enough duration of training, memory of the initial conditions should be erased and the final learning outcome should be the same with or without pre-training.
In \autoref{fig:longterm} we show the result of longer training durations on the three models that were not ruled out on the basis of the initial learning rate, the serial, cascade and nonuniform multistate models.
For all of these models, we see that the wild-type models eventually catch up and overtake the \KO\.
The initial learning enhancement is a transient effect, and at later times it turns into a learning impairment.
This is consistent with an extrapolation of the trend in \datafig.

@@

We see that, for all parameter values, such models do not reproduce the observed learning impairment with enhanced plasticity.
The normal baseline oculomotor performance must have some other explanation, such as those outlined in \compfig[].
